\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-25 NAR_DRAFT.tex
fix intro citations

Many methods have been developed for calling somatic mutations from NGS data.
The earliest widely used somatic variant callers developed specifically for tumors, MuTect1~\cite{Cibulskis2013} and Varscan2~\cite{Koboldt2012}, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants and set a threshold score designed to balance sensitivity and specificity.
Subsequent research gave rise to a number of alternate strategies, including haplotype-based calling \citep{Garrison2012},
joint genotype analysis (SomaticSniper, JointSNVMix2, Seurat, and CaVEMan,MuClone) \cite{Larson2012,Roth2012a,Christoforides2013,Jones2016,Dorri2019}, allele frequency-based analysis (Strelka, MuTect, LoFreq, EBCall, deepSNV, LoLoPicker, and MuSE) \citep{Saunders2012,Wilm2012,Shiraishi2013b,Gerstung2012,Carrot-Zhang2017,Fan2016}, and ensemble and deep learning methods (MutationSeq, BAYSIC, SomaticSeq, and SNooPer) \citep{Ding2012,Cantarel2014,Fang2015,Spinella2016}.
\rngcomment{Connect citations to individual tools. (See MuTect1 and Varscan2 examples earlier in paragraph.}
These methods vary in their complexity and specific focus.
But they all implicitly or explicitly assume that the rate of mutation is uniform across the genome.

>>>>

Many methods have been developed for calling somatic mutations from NGS data.
The earliest widely used somatic variant callers developed specifically for tumors, MuTect1~\cite{Cibulskis2013} and Varscan2~\cite{Koboldt2012}, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants and set a threshold score designed to balance sensitivity and specificity.
Subsequent research gave rise to a number of alternate strategies, including haplotype-based calling \citep{Garrison2012},
joint genotype analysis (SomaticSniper \citep{Larson2012}, JointSNVMix2 \citep{Roth2012a}, Seurat \citep{Christoforides2013}, CaVEMan \citep{Jones2016}, and MuClone \citep{Dorri2019}), allele frequency-based analysis (Strelka \citep{Saunders2012}, LoFreq \citep{Wilm2012}, EBCall \citep{Shiraishi2013b}, deepSNV \citep{Gerstung2012}, LoLoPicker \citep{Carrot-Zhang2017}, and MuSE \citep{Fan2016}), and ensemble and deep learning methods (MutationSeq \citep{Ding2012}, BAYSIC \citep{Cantarel2014}, SomaticSeq \citep{Fang2015}, and SNooPer \citep{Spinella2016}).
These methods vary in their complexity and specific focus.
But they all implicitly or explicitly assume that the rate of mutation is uniform across the genome.
